Genetic draft and valley crossing

### 2019-05-24 ms/valley_crossing.tex
adjust fig 2

Valley crossing in asexual populations is increasingly well understood \citep{weissman_2009}.
If the number of mutants produced every generation is high, as is the case in large populations with high mutation rates, a complex adaptation requiring multiple mutations that is destined to fix will eventually be generated \emph{de novo}.
This is sometimes called deterministic fixation of the multiple mutant \citep{weissman_2009}.
In small populations, neutral or deleterious intermediate mutants generated by wild type individuals can drift to fixation. Further mutations necessary for the adaptation can then arise on intermediate mutant backgrounds and fix due to positive selection.
This is the sequential fixation regime \citep{weissman_2009}.
A third possibility occurs for intermediate population sizes wherein wild type individuals generate transient mutant subpopulations or ``bubbles".
These bubbles are ordinarily doomed to extinction due to drift, but additional mutations in a lucky bubble may generate a complex adaptation that sweeps to fixation; this process has been referred to as ``tunneling" \citep{iwasa_2004, weissman_2009}.
On its own, valley crossing becomes more likely as the population size increases \citep{weissman_2009}. However, recent work suggests that valley crossing relative to the fixation of a simple beneficial mutation is least likely at intermediate population sizes in which tunneling occurs \citet{ochs_2015}. This suggests that tunneling may be the most difficult mode of valley crossing. 
Recombination also affects the rate of valley crossing.
Infrequent recombination decreases valley crossing times \citep{weissman_2010}.
At high values of the recombination rate, however, individuals carrying the full complex adaptation outcross with the wild type and produce deleterious intermediates that retard the growth of the multiple mutant.
An additional effect that can heighten the rate at which valley crossing occurs is population subdivision \citep{Bitbol:Schwab:2014}: the population size within a subdivision is smaller, meaning that selection against potentially harmful intermediates is relaxed.

Previous studies of valley crossing focus primarily on populations where all individuals are equal in fitness except for the focal loci, i.e., the loci at which the individual mutations comprising the complex adaptation segregate.
This is tantamount to assuming that, if there is variation in the background fitness of the population, it is negligible.
In such populations, genetic drift governs the fate of neutral alleles, as well as the behavior of deleterious or beneficial alleles where they are rare (close to frequency zero) or common (close to frequency one).
Alleles whose dynamics are primarily determined by genetic drift can be effectively modeled by a diffusion approximation \citep{Wright:1945,Kimura:1955,Kimura:1957}, and the ancestry of a population in the absence of selection is described well by the classic Kingman coalescent \citep{Kingman:1982}, where pairs of branches in the genealogy merge back in time until the most recent common ancestor.
When population sizes are very large or the fitness variation in the population is substantial, however, the behavior of genetic variation is governed more by genetic \emph{draft} than by genetic drift \citep{gillespie_2000, gillespie_2001, masel_2011, neher_shraiman_2011}.
Whereas drift is the effect of imperfect sampling from generation to generation, draft is the sum of hitchhiking effects due to selection on sites linked to the focal loci.
The standard technique of defining a reduced ``effective population size" fails to capture several fundamental qualitative differences between drift and draft.
For example, new mutations in a drafting population can rapidly be brought to high frequency, and the variance in offspring number almost diverges \citep{neher_shraiman_2011}.
Drafting populations commonly have genealogies in which more than two branches merge at once (or almost at once) and that are better described by alternative coalescent processes such as the Bolthausen-Sznitman coalescent \citep{neher_hallatschek_2013, brunet_2007, schweinsberg_2017}.
Additionally under genetic draft, the frequency spectrum of neutral alleles is non-monotonic, with a marked uptick near frequency one--or, equivalently, a depression in intermediate frequency alleles, which are quickly swept out of the population \citep{neher_shraiman_2011,kosheleva_2013,neher_hallatschek_2013}.
Only a handful of individuals near the ``nose" of the fitness distribution are likely to persist, and they give rise to the bulk of the future population and carry linked alleles with them.
The parameter that determines whether drift or draft is more important is the product of the population size $N$ and the standard deviation in fitness $\sigma$ \citep{neher_hallatschek_2013}.

It is increasingly realized that draft may play a critical role in shaping genetic diversity, especially in many microbial species, where population sizes can be large but ``effective population sizes" are many orders of magnitude smaller \citep{masel_2011}.
The possibility that neutral variants are affected more strongly by selection at linked sites than by genetic drift, even in organisms like \emph{Drosophila}, is one possible explanation for the ``paradox of variation", the fact that genetic diversity and population size often do not scale linearly \citep{gillespie_2000, gillespie_2001, neher_kessinger_2013, corbett-detig_2015}.
Draft likewise may have profound effects on the evolution of complex adaptations. Preliminary evidence for this comes from
\citet{neher_shraiman_2011}, who explored how draft affects valley crossing via stochastic tunneling by calculating the mutant bubble size distribution in a rapidly adapting facultatively sexual population.
They found that, compared to drift, draft generally causes the distribution of bubble sizes to drop off more quickly, meaning that bubbles tend to be smaller overall: however, as the mutant lineage's fitness is determined not only by a mutation's intrinsic fitness effect but also the genetic background on which it appears, the fitness effect of a focal mutation plays less of a role in capping the size of the lineage.
Thus, while draft may make valley crossing via tunneling more difficult, the total effect isn't immediately obvious.
Here, we extend these previous results by studying how genetic draft in asexual populations affects the rate of crossing fitness valleys across a range of population sizes.

The complicating factor is that, in an asexual population, the dynamics are governed almost entirely by what happens in the nose of the fitness distribution, and these dynamics can be highly stochastic.
In previous approaches, these stochastic effects were smoothed due to the presence of recombination \citep{neher_shraiman_2011, neher_shraiman_2009}: the critical nose dynamics are still important in a sexual population, but the fitness distribution is more Gaussian overall.
We therefore must focus primarily on simulation approaches, as analytical solutions for the behavior of a mutation in the nose of an asexual population are very difficult to obtain.
We expand on the work of \citet{neher_shraiman_2011} by considering not only stochastic tunneling but also sequential fixation; in fact, we find that, over a significant chunk of parameter space, stochastic tunneling effectively \emph{does not happen} in drafting populations, and sequential fixation appears to be the dominant mode of valley crossing.
We also show that fitness valley crossing occurs at an overall lower rate in rapidly adapting populations.
This is consistent with the fact that clonal interference and genetic background effects retard the growth of fit mutant lineages, slowing the expansion of a beneficial mutation.
In addition, we confirm that in rapidly adapting populations, adaptive fixations of alleles are more likely to be complex adaptations involving a fitness valley than they are in populations where genetic drift is the primary force shaping genetic variation at linked sites; essentially, genetic draft maintains the linkage that makes it possible to leap across fitness landscapes rather than adapt primarily by climbing to local peaks.
These observations add to the growing intuition that complex adaptations that involve evolution across fitness valleys may not be a surprising result of the evolutionary process, but rather an expected one.

@@

We further assume that the large number of background loci of weak effect contribute to a constant background fitness variance $\sigma^2$.
Let $x$ be the background fitness of a lineage.
If the number of background loci is large, the distribution of background fitnesses $f(x)$ will generally be a series of (Dirac) $\delta$ functions, but its ensemble average will be roughly Gaussian \citep{good_desai_2014}: $f(x) = \frac{1}{\sqrt{2\pi}\sigma} \exp (\frac{(x-\bar{x})^2}{2\sigma^2})$.
By Fisher's ``fundamental theorem", the mean fitness $\bar{x}$ will advance due to selection at a rate equal to $\sigma^2$: the total rate of advance will depend also on contributions from drift and the average effect of mutations.

@@

\citet{weissman_2009} divided the possible modes of asexual valley crossing into several distinct regimes.
We classify them, broadly, as \emph{sequential fixation}, \emph{deterministic fixation}, and \emph{tunneling}.
What follows is a brief outline of the process of valley crossing in each of these regimes, as well as how these are likely to differ in our model.
We then present results for valley crossing times using forward-time simulations across a range of parameter values that includes each of these major regimes.

The \emph{sequential fixation} regime arises at small population sizes.
The wait time $\mathbb{E}\left[ T\right]$ can be partitioned into two components: the time $\tau_1$ until an intermediate mutation arises \emph{that is destined to fix} and the time $\tau_2$ until the complex adaptation arises and is likewise destined to fix.
$\tau_1$ is one divided by the rate $N\mu_1 p$ at which intermediate mutants arise that are destined to fix: on average, $N\mu_1$ mutants appear every generation, and the diffusion approximation yields $p = (e^\delta - 1)/(e^{N\delta} - 1)$.
$\tau_2$, similarly, is approximately one divided by $N\mu_2 s$ and in general will be dwarfed by $\tau_1$.
When genetic draft is more important than genetic drift, $p$ in $\tau_1$ takes on a much more complicated form similar to that in \citet{good_desai_2014}: in general, the dependence on $\delta$ is lessened provided that $\sigma \gtrsim \delta$, as deleterious mutations that arise on significantly fitter-than-average backgrounds can be destined to fix.
$\tau_2$ likewise has lessened dependence on $s$, as a complex adaptation that does \emph{not} arise on a fitter-than-average background lineage may be doomed to extinction.

There are three components to the process of stochastic tunneling.
First, a single mutant lineage must appear, which occurs with rate $N\mu$.
If this lineage arises at time $t_0$ and persists until time $T$, then it gives rise to $W = \int_{t_0}^T n(t) dt$ total individuals before going extinct, where $n(t)$ is the number of mutant individuals extant at time $t$.
As in the introduction, we refer to such a short-lived mutant lineage, and all the individuals generated within that lineage, as a mutant ``bubble": the time integrated number of individuals $W$ is the ``weight" of the bubble.
If it establishes, the complex adaptation will quickly sweep to fixation.
The first two steps can be folded in together and considered as one process, so that what is relevant is the expected time until a single mutant lineage arises that is destined to produce a successful double mutant.
In this way, $\mathbb{E}\left[ T \right] = \mathbb{E} \left[ T_0 + T_1 \right]$, where $T_0$ is the time to the first successful bubble--the first that produces a double mutant that is destined for fixation--and $T_1$ is the time required for the double mutant to sweep.
We assume that $T_0 \gg T_1$, so that the wait time is dominated by the first term: in the drift case, the time required for $T_1$ to establish will be of order $1/s$ \citep{desai_fisher_2007}, and the sweep time will be of order $1/s \log (Ns)$, but $T_0$ can be many orders of magnitude greater.

@@

To analyze our model, we perform forward simulations using the custom Julia module \texttt{WaveFit}, which implements a modified Wright-Fisher model.
The simulation code is available via \texttt{GitHub}.
In \texttt{WaveFit}, individuals are organized by ``clones" (sets of individuals with the same genome): each clone gives rise to a Poisson distributed number of offspring, with the mean dependent on the relative fitness $x-\bar{x}$.
The mean offspring number is further adjusted to keep the population roughly at a pre-specified carrying capacity $K$.
In addition, each individual receives a Poisson distributed number of mutations: if $L$ is the number of sites in the genome and $U$ is the (background) per-site mutation rate, then $UL$ is the mean number of mutations per individual per generation (and is a parameter in our simulations).
In effect, this is an ``infinite sites" model \citep{Kimura:1971,Watterson:1975}.
All background loci are assigned a common fitness effect $\epsilon$, which is scaled up or down each generation so as to keep the background fitness standard deviation $\sigma$ constant.
This rescaling means that the assigned initial value of $\epsilon$ turns out not to matter, even if it is negative or even if both deleterious and beneficial background mutations are permitted.
That feature of our model may seem counterintuitive, but previous work has found that populations experiencing strong background selection at many loci of weak effect (see, e.g., \citet{cvijovic_2015}) exhibit dynamics very similar to rapidly adapting populations \citep{desai_fisher_2007, neher_shraiman_2011}.

@@

\begin{figure*}[t]
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=\textwidth]{Figures/julia_weissman_5a.pdf}
%\label{fig:weissmana}
\end{subfigure}
\begin{subfigure}[b]{0.48\textwidth}
\includegraphics[width=\textwidth]{Figures/julia_weissman_5b.pdf}
%\label{fig:weissmanb}
\end{subfigure}
\begin{subfigure}[c]{0.48\textwidth}
\includegraphics[width=\textwidth]{Figures/julia_weissman_5c.pdf}
%\label{fig:weissmanc}
\end{subfigure}
\begin{subfigure}[d]{0.48\textwidth}
\includegraphics[width=\textwidth]{Figures/julia_weissman_5d.pdf}
%\label{fig:weissmand}
\end{subfigure}
\caption{Valley crossing times across a range of parameter values mirroring those of \citet{weissman_2009}. In each case, we set $s = 0.1$. In figures (a) and (b), the population size is varied: increasing the size causes valley crossing to transition from tunneling to semi-deterministic fixation. In (a), $\mu_1 = 10^{-5}$, $\mu_2 = 10^{-4}$, and $\delta = 2 \times 10^{-4}$, so that the valley depth is effectively neutral; in (b), $\mu_1 = 10^{-5}$, $\mu_2 = 10^{-6}$, and $\delta = 7 \times 10^{-3}$. In figure (c), we set $N = 10^5$, $\mu_1 = 10^{-6}$, and $\mu_2 = 4 \times 10^{-5}$. The valley depth $\delta$ is varied: at negative $\delta$, our simulations closely match the expectation for beneficial mutations. In figure (d), the intermediate mutation rate $\mu_2$ is varied, with $N = 10^{5}$, $\mu_1 = 10^{-6}$, and $\delta = 3 \times 10^{-3}$. Dotted lines represent the theoretical predictions from \citet{weissman_2009}.}
\label{fig:weissman}
\end{figure*}

>>>>

Valley crossing in asexual populations is increasingly well understood \citep{weissman_2009}.
If the number of mutants produced every generation is high, as is the case in large populations with high mutation rates, a complex adaptation requiring multiple mutations that is destined to fix will eventually be generated \emph{de novo}.
This is sometimes called deterministic fixation of the multiple mutant \citep{weissman_2009}.
In small populations, neutral or deleterious intermediate mutants generated by wild type individuals can drift to fixation. Further mutations necessary for the adaptation can then arise on intermediate mutant backgrounds and fix due to positive selection.
This is the sequential fixation regime \citep{weissman_2009}.
A third possibility occurs for intermediate population sizes wherein wild type individuals generate transient mutant subpopulations or ``bubbles".
These bubbles are ordinarily doomed to extinction due to drift, but additional mutations in a lucky bubble may generate a complex adaptation that sweeps to fixation; this process has been referred to as ``tunneling" \citep{iwasa_2004, weissman_2009}.
On its own, valley crossing becomes more likely as the population size increases \citep{weissman_2009}.
However, recent work suggests that valley crossing relative to the fixation of a simple beneficial mutation is least likely at intermediate population sizes in which tunneling occurs \citet{ochs_2015}.
This suggests that tunneling may be the most difficult mode of valley crossing. 
Recombination also affects the rate of valley crossing.
Infrequent recombination decreases valley crossing times \citep{weissman_2010}.
At high values of the recombination rate, however, individuals carrying the full complex adaptation outcross with the wild type and produce deleterious intermediates that retard the growth of the multiple mutant.
An additional effect that can heighten the rate at which valley crossing occurs is population subdivision \citep{Bitbol:Schwab:2014}: the population size within a subdivision is smaller, meaning that selection against potentially harmful intermediates is relaxed.
Intermediates can then drift to high frequency or even fixation, increasing the rate at which complex adaptations arise.

Previous studies of valley crossing focus primarily on populations where all individuals are equal in fitness except for the focal loci, i.e., the loci at which the individual mutations comprising the complex adaptation segregate.
This is tantamount to assuming that variation in the background fitness of the population is negligible.
In such populations, genetic drift governs the fate of neutral alleles, as well as the behavior of deleterious or beneficial alleles where they are rare (close to frequency zero) or common (close to frequency one).
Alleles whose dynamics are primarily determined by genetic drift can be effectively modeled by a diffusion approximation \citep{Wright:1945,Kimura:1955,Kimura:1957}, and the ancestry of a population in the absence of selection is described well by the classic Kingman coalescent \citep{Kingman:1982}, where pairs of branches in the genealogy merge back in time until the most recent common ancestor.
When population sizes are very large or the fitness variation in the population is substantial, however, the behavior of genetic variation is governed more by genetic \emph{draft} than by genetic drift \citep{gillespie_2000, gillespie_2001, masel_2011, neher_shraiman_2011}.
Broadly speaking, drift results from the fact that each generation is imperfectly sampled from the previous generation.
Draft, on the other hand, is the sum of hitchhiking effects due to selection on sites linked to the focal loci.
The standard technique of defining a reduced ``effective population size" fails to capture several fundamental qualitative differences between drift and draft.
For example, new mutations in a drafting population can rapidly be brought to high frequency, and the variance in offspring number almost diverges \citep{neher_shraiman_2011}.
Drafting populations commonly have genealogies in which more than two branches merge at once (or almost at once) and that are better described by alternative coalescent processes such as the Bolthausen-Sznitman coalescent \citep{neher_hallatschek_2013, brunet_2007, schweinsberg_2017}.
Additionally under genetic draft, the frequency spectrum of neutral alleles is non-monotonic, with a marked uptick near frequency one--or, equivalently, a depression in intermediate frequency alleles, which are quickly swept out of the population \citep{neher_shraiman_2011,kosheleva_2013,neher_hallatschek_2013}.
Only a handful of individuals near the ``nose" of the fitness distribution are likely to persist, and they give rise to the bulk of the future population and carry linked alleles with them.
The parameter that determines whether drift or draft is more important is the product of the population size $N$ and the standard deviation in fitness $\sigma$ \citep{neher_hallatschek_2013}.

It is increasingly realized that draft may play a critical role in shaping genetic diversity, especially in many microbial species, where population sizes can be large but ``effective population sizes" are many orders of magnitude smaller \citep{masel_2011}.
The possibility that neutral variants are affected more strongly by selection at linked sites than by genetic drift, even in organisms like \emph{Drosophila}, is one possible explanation for the ``paradox of variation", the fact that genetic diversity and population size often do not scale linearly \citep{gillespie_2000, gillespie_2001, neher_kessinger_2013, corbett-detig_2015}.
Draft likewise may have profound effects on the evolution of complex adaptations.
Preliminary evidence for this comes from \citet{neher_shraiman_2011}, who explored how draft affects valley crossing via stochastic tunneling by calculating the mutant bubble size distribution in a rapidly adapting facultatively sexual population.
They found that, compared to drift, draft generally causes the distribution of bubble sizes to drop off more quickly, meaning that bubbles tend to be smaller overall: however, as the mutant lineage's fitness is determined not only by a mutation's intrinsic fitness effect but also the genetic background on which it appears, the fitness effect of a focal mutation plays less of a role in capping the size of the lineage.
Thus, while draft may make valley crossing via tunneling more difficult, the total effect is not immediately obvious.
Here, we extend these results by studying how genetic draft in asexual populations affects the rate of crossing fitness valleys across a range of population sizes, including ones in which tunneling is not the dominant mode of valley crossing.

The complicating factor is that, in an asexual population, the dynamics are governed almost entirely by what happens in the nose of the fitness distribution, and these dynamics can be highly stochastic.
In previous approaches, these stochastic effects were smoothed due to the presence of recombination \citep{neher_shraiman_2011, neher_shraiman_2009}: the critical nose dynamics are still important in a sexual population, but the fitness distribution is more Gaussian overall.
Moreover, a full understanding of the behavior of complex adaptations requires that one consider the dynamics of individual lineages.
The equations that describe said dynamics, however, are very difficult to solve; \citet{neher_shraiman_2011} were able to circumvent this issue by carefully choosing a recombination model that allowed them to avoid the need to directly solve these equations.
We therefore must focus primarily on simulation approaches, as analytical solutions for the behavior of a mutation in the nose of an asexual population are very difficult to obtain.
We expand on the work of \citet{neher_shraiman_2011} by considering not only stochastic tunneling but also sequential fixation; in fact, we find that, over a significant chunk of parameter space, stochastic tunneling effectively \emph{does not happen} in drafting populations, and sequential fixation appears to be the dominant mode of valley crossing.
We also show that fitness valley crossing occurs at an overall lower rate in rapidly adapting populations.
This is consistent with the fact that clonal interference and genetic background effects retard the growth of fit mutant lineages, slowing the expansion of a beneficial mutation.
In addition, we confirm that in rapidly adapting populations, adaptive fixations of alleles are more likely to be complex adaptations involving a fitness valley than they are in populations where genetic drift is the primary force shaping genetic variation at linked sites; essentially, genetic draft maintains the linkage that makes it possible to leap across fitness landscapes rather than adapt primarily by climbing to local peaks.
These observations add to the growing intuition that complex adaptations that involve evolution across fitness valleys may not be a surprising result of the evolutionary process but rather an expected one.

@@

We further assume that the large number of background loci of weak effect contribute to a constant background fitness variance $\sigma^2$.
Let $x$ be the background fitness of a lineage.
If the number of background loci is large, the distribution of background fitnesses $f(x)$ will generally be a series of (Dirac) $\delta$ functions, but its ensemble average will be roughly Gaussian \citep{good_2014}: $f(x) \approx \frac{1}{\sqrt{2\pi}\sigma} \exp (\frac{(x-\bar{x})^2}{2\sigma^2})$.
By Fisher's ``fundamental theorem", the mean fitness $\bar{x}$ will advance due to selection at a rate equal to $\sigma^2$: the total rate of advance will depend also on contributions from drift and the average effect of mutations.

@@

\citet{weissman_2009} divided the possible modes of asexual valley crossing into several distinct regimes.
We classify them, broadly, as \emph{sequential fixation}, \emph{deterministic fixation}, and \emph{stochastic tunneling}.
What follows is a brief outline of the process of valley crossing in each of these regimes, as well as how these are likely to differ in our model.
We then present results for valley crossing times using forward-time simulations across a range of parameter values that includes each of these major regimes.

The \emph{sequential fixation} regime arises at small population sizes.
The wait time $\mathbb{E}\left[ T\right]$ can be partitioned into two components: the time $\tau_1$ until an intermediate mutation arises \emph{that is destined to fix} and the time $\tau_2$ until the complex adaptation arises and is likewise destined to fix.
$\tau_1$ is one divided by the rate $N\mu_1 p$ at which intermediate mutants arise that are destined to fix: on average, $N\mu_1$ mutants appear every generation, and the diffusion approximation yields $p = (e^\delta - 1)/(e^{N\delta} - 1)$.
$\tau_2$, similarly, is approximately one divided by $N\mu_2 s$ and in general will be dwarfed by $\tau_1$.
When genetic draft is more important than genetic drift, $p$ in $\tau_1$ takes on a much more complicated form similar to that in \citet{good_2014}: in general, the dependence on $\delta$ is lessened provided that $\sigma \gtrsim \delta$, as deleterious mutations that arise on significantly fitter than average backgrounds can be destined to fix.
$\tau_2$ likewise has lessened dependence on $s$, as a complex adaptation that does \emph{not} arise on a fitter-than-average background lineage may be doomed to extinction.

The process of \emph{deterministic fixation} occurs when the number of single mutants introduced every generation is sufficiently large: $N\mu_1 \gg 1$.
One possibility is that the intermediate exists in what is roughly mutation-selection balance, so that it persists at a low frequency.
Lucky complex adaptations can then arise on this background of roughly constant intermediate influx, then sweep to fixation.
This is known as \emph{semi-deterministic tunneling}; we class it under the ``deterministic" umbrella because it is not necessary to consider the stochastic dynamics of individual deleterious mutant lineages.
Switching from drift to draft will affect the likelihood that a particular complex adaptation will sweep to fixation, but it seems unlikely to affect the mutation-selection balance dynamics for intermediates.
The strictest ``deterministic fixation" case occurs when $N \gg 2s/\pi \mu_1 \mu_2$.
In this case, multiple complex adaptations are likely to arise on independent backgrounds, and there is no guarantee that the first one to reach significant frequency will sweep through the population on its own.
In the draft case, we expect the fixation time to be increased due to the blunting effect of the genetic background.
However, note that the $N \gg 2s/\pi \mu_1 \mu_2$ cutoff is generally very large: for $\mu_1 = \mu_2 = 10^{-4}$ and $s = 0.1$, we have $N \approx 10^7$, a size that will prove difficult to probe with our simulation model.

Finally, we consider the intermediate case: \emph{stochastic tunneling}.
This process has three major components.
First, a single mutant lineage must appear, which occurs with rate $N\mu$.
If this lineage arises at time $t_0$ and persists until time $T$, then it gives rise to $W = \int_{t_0}^T n(t) dt$ total individuals before going extinct, where $n(t)$ is the number of mutant individuals extant at time $t$.
As in the introduction, we refer to such a short-lived mutant lineage, and all the individuals generated within that lineage, as a mutant ``bubble": the time integrated number of individuals $W$ is the ``weight" of the bubble.
If it establishes, the complex adaptation will quickly sweep to fixation.
The first two steps can be folded in together and considered as one process, so that what is relevant is the expected time until a single mutant lineage arises that is destined to produce a successful double mutant.
In this way, $\mathbb{E}\left[ T \right] = \mathbb{E} \left[ T_0 + T_1 \right]$, where $T_0$ is the time to the first successful bubble--the first that produces a double mutant that is destined for fixation--and $T_1$ is the time required for the double mutant to sweep.
We assume that $T_0 \gg T_1$, so that the wait time is dominated by the first term: in the drift case, the time required for $T_1$ to establish will be of order $1/s$ \citep{desai_fisher_2007}, and the sweep time will be of order $1/s \log (Ns)$, but $T_0$ can be many orders of magnitude greater.

@@

To analyze our model, we perform forward simulations using the custom Julia module \texttt{WaveFit}, which implements a modified Wright-Fisher model.
The simulation code is available via \texttt{GitHub}.
In \texttt{WaveFit}, individuals are organized by ``clones" (sets of individuals with the same genome): each clone gives rise to a Poisson distributed number of offspring, with the mean dependent on the relative fitness $x-\bar{x}$.
The mean offspring number is further adjusted to keep the population roughly at a pre-specified carrying capacity $K$.
In addition, each individual receives a Poisson distributed number of mutations: if $L$ is the number of sites in the genome and $U$ is the (background) per-site mutation rate, then $UL$ is the mean number of mutations per individual per generation (and is a parameter in our simulations).
In effect, this is an ``infinite sites" model \citep{Kimura:1971,Watterson:1975}.
All background loci are assigned a common fitness effect $\epsilon$, which is scaled up or down each generation so as to keep the background fitness standard deviation $\sigma$ constant.
This rescaling means that the assigned initial value of $\epsilon$ turns out not to matter, even if it is negative or even if both deleterious and beneficial background mutations are permitted.
That feature of our model may seem counterintuitive, but previous work has found that populations experiencing strong background selection at many loci of weak effect (see, e.g., \citet{cvijovic_2018}) exhibit dynamics very similar to rapidly adapting populations \citep{desai_fisher_2007, neher_shraiman_2011}.

@@

\begin{figure*}[!t]
  \begin{tabular}{r@{\hspace{-1ex}}r}
\includegraphics[width=0.48\textwidth]{Figures/julia_weissman_5a.pdf} &
%\label{fig:weissmana}
\includegraphics[width=0.48\textwidth]{Figures/julia_weissman_5b.pdf}\\[-3ex]
%\label{fig:weissmanb}
\includegraphics[width=0.48\textwidth]{Figures/julia_weissman_5c.pdf} &
%\label{fig:weissmanc}
\includegraphics[width=0.48\textwidth]{Figures/julia_weissman_5d.pdf}
% \label{fig:weissmand}
  \end{tabular}
  \caption{Valley crossing times across a range of parameter values mirroring those of \citet{weissman_2009}. In each case, we set $s = 0.1$. In figures (a) and (b), the population size is varied: increasing the size causes valley crossing to transition from tunneling to semi-deterministic fixation. In (a), $\mu_1 = 10^{-5}$, $\mu_2 = 10^{-4}$, and $\delta = 2 \times 10^{-4}$, so that the valley depth is effectively neutral; in (b), $\mu_1 = 10^{-5}$, $\mu_2 = 10^{-6}$, and $\delta = 7 \times 10^{-3}$. In figure (c), we set $N = 10^5$, $\mu_1 = 10^{-6}$, and $\mu_2 = 4 \times 10^{-5}$. The valley depth $\delta$ is varied: at negative $\delta$, our simulations closely match the expectation for beneficial mutations. In figure (d), the intermediate mutation rate $\mu_2$ is varied, with $N = 10^{5}$, $\mu_1 = 10^{-6}$, and $\delta = 3 \times 10^{-3}$. Dotted lines represent the theoretical predictions from \citet{weissman_2009}.}
\label{fig:weissman}
\end{figure*}
